### 2024-11-02 to_do_list/To_do_list.qmd

**Format:** pdf

### **Person A: Data Preparation & Pollster Analysis**

- **Task 1: Data Documentation & Review**
  - Review and finalize data cleaning documentation, ensuring reproducibility and clarity.
  - Provide a summary of any challenges encountered during data cleaning, along with adjustments made to the data.

- **Task 2: Model Explanation & Assumptions**
  - Define all model components, including notation and plain-English explanations.
  - Discuss and document model assumptions, limitations, and potential issues.
  - Validate model performance using test/train splits, RMSE calculations, or sensitivity analyses.

- **Task 3: Data Section Write-Up (Assistance with Visualizations)**
  - Collaborate with Person C to finalize the data section, focusing on key visualizations and summary statistics.
  - Ensure all visualizations and statistics are clearly presented.

---

### **Person B: Model Building, Analysis, & Validation**

- **Task 1: Pollster Deep-Dive for Appendix A**
  - Choose a pollster from the data sample and research their methodology in-depth, focusing on survey design (population, frame, sample recruitment, etc.).
  - Write an overview of the pollster’s strengths, weaknesses, and key features.

- **Task 2: Initial Discussion Write-Up**
  - Draft the results and discussion section based on model outcomes, highlighting key findings and their relevance to the election forecast.
  - Outline potential improvements or next steps based on the model’s performance and limitations.

- **Task 3: Results Section Write-Up**
  - Develop tables, graphs, and visualizations to present model findings, ensuring clear interpretation.

---

### **Person C: Documentation, Workflow, & Final Integration**

- **Task 1: Results Section Write-Up**
  - Create tables, graphs, and visualizations to clearly present the model findings.
  - Interpret model results and include any relevant performance metrics (e.g., RMSE, test/train split).

- **Task 2: Reproducible Workflow & Documentation**
  - Organize the GitHub repo with folders (`data/`, `scripts/`, `figures/`, `docs/`).
  - Finalize the README file detailing project structure, data sources, and model.

- **Task 2: Idealized Survey Design for Appendix B**
  - Design an ideal survey methodology for forecasting the election on a $100K budget.
  - Complete the survey on Google Forms, including the link and a PDF export in Appendix B.

- **Task 4: Final Paper Integration & Formatting**
  - Compile and format all sections into a cohesive document by **November 2nd**. Add cross-references to figures, tables, and equations.

---

# Timeline Updates

### **Week 3: October 21st - October 22nd**
- **Person A:** Complete Pollster Deep-Dive for Appendix A and collaborate on the Data Section Write-Up.

### **Week 4: October 23rd - October 29th**
- **Person B:** Complete model development, analysis, and initial validation.

### **Week 5: October 30th - November 4th**
- **Person C:** Finalize integration, quality checks, and prepare for submission.